---
title: "Estimating standard errors clustered by firm and time (Petersen, RFS 2009)"
output: html_document
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
      font-family: "Times New Roman", Times, serif;
  }
td {  /* Table  */
  font-size: 20px;
}
h1.title {
  font-size: 24px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 22px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 22px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 16px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 16px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reference

- https://academic.oup.com/rfs/article-abstract/22/1/435/1585940?redirectedFrom=fulltext
- https://www.kellogg.northwestern.edu/faculty/petersen/htm/papers/standarderror.html


This note demonstrates how to calculate standard errors clustered by firm and time in R project, following Petersen (RFS, 2009).

```{r, message=FALSE}
rm(list=ls());
library(sandwich);
library(plm);
library(lmtest);
library(stargazer);
```

- Data source: http://www.kellogg.northwestern.edu/faculty/petersen/htm/papers/se/test_data.txt

```{r}
df <- read.csv(file="01_Data/test_data.txt")
summary(df)
```
## OLS

```{r}
reg0 = lm(y~x, data=df)
ols = coeftest(reg0)
ols
```

## OLS with White Standard Errors

- https://en.wikipedia.org/wiki/Heteroscedasticity-consistent_standard_errors

```{r}
reg0 = lm(y~x, data=df)
white = coeftest(reg0, vcov = function(x) vcovHC(x, method="white1", type="HC1"))
white
```

## OLS with Newey-West
- https://en.wikipedia.org/wiki/Newey%E2%80%93West_estimator

```{r}
reg0 = lm(y~x, data=df)
newey = coeftest(reg0, vcov = NeweyWest(reg0))
newey
```

## OLS clustered by Firm

```{r}
reg1 = plm(y~x, model="pooling", index=c("firm", "year"), data=df)
cluster.firm = coeftest(reg1, vcov = function(x) vcovHC(x, cluster="group", type="HC1"))
cluster.firm
```

## OLS clustered by Time (Year), R Project Setup

```{r}
reg1 = plm(y~x, model="pooling", index=c("firm", "year"), data=df)
cluster.time.R = coeftest(reg1, vcov = function(x) vcovHC(x, cluster="time", type="HC1"))
cluster.time.R
```

## OLS clustered by Time (Year), STATA Setup

```{r}
reg1 = plm(y~x, model="pooling", index=c("firm", "year"), data=df)
cluster.time.stata = coeftest(reg1, vcov = function(x) vcovHC(x, method=c("arellano"), type=c("sss"), cluster = c("time"))) 

cluster.time.stata
```

## OLS clustered by Firm and Year - STATA Default

```{r}
vcovDC = function(x, ...){
    vcovHC(x, cluster="group", ...) + 
      vcovHC(x, method=c("arellano"), type=c("sss"),cluster = c("time"), ...) - 
      vcovHC(x, method="white1", ...)
  }
reg1 = plm(y~x, model="pooling", index=c("firm", "year"), data=df)
cluster.double = coeftest(reg1, vcov = function(x) vcovDC(x))
cluster.double
```

# Fama-MacBeth Regression

- https://en.wikipedia.org/wiki/Fama%E2%80%93MacBeth_regression- 


```{r}
fmb = pmg(y~x, df, index=c("year","firm"))
FMB = coeftest(fmb)
FMB
```

```{r results = "asis"}
stargazer::stargazer(ols,white,newey,cluster.time.R, cluster.time.stata,cluster.firm,cluster.double,FMB,type = "html", column.labels = c("OLS", "White", "NW","C.T(R)","C.T(S)", "C.F", "C.FT","FM"),
 title = "Adjusting Standard Errors")
```




# Petersen's Results

- https://www.kellogg.northwestern.edu/faculty/petersen/htm/papers/se/test_data.htm


